---
title: "glowworms_data_reproduce"
format: html
---

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
source(here::here("data-raw/data_reproduce_utiles.R"))
```

# Define Global Parameters

We define basic variables used throughout the vignette:

```{r echo=TRUE, eval=TRUE}
# global variables
email <- "your_email@email.com"
organism_name <- "glowworms"
organism_title <- "Glowworms"
taxa     <- "Arachnocampa"
start    <- 2014
end      <- 2024
```

## Step 1: Download and Prepare Occurrence Data

```{r eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
organism <- obs_data(taxa, start, end, your_email=email) |> polished_data()
```

> This downloads ALA data and performs standard cleaning: renaming columns and removing `NA`s.

## Step 2: Save Cleaned Dataset

```{r eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
mysave(organism_name, organism)
```

This saves the object to the `/data` folder with the name `"glowworms.rda"`.

## Step 3: Attach Nearest Weather Station

```{r, echo=TRUE, message=FALSE, eval=FALSE}
attach_nearest_station(organism_name)
```

> Weather data is restricted to the top 3 stations to keep it lightweight.

## Step 4: Load and Explore the Data

```{r, echo=TRUE, eval=FALSE}
load("../data/glowworms.rda")
glimpse(glowworms)
```

## Step 5: Save Raw Version for Further Analysis

```{r "request data ALA", eval=FALSE, message=FALSE, warning=FALSE}
organism_raw <- obs_data(taxa, start, end, your_email=email) |> polished_data()

mysave_raw(organism_name, organism_raw)
```

To load it:

```{r , echo=TRUE, eval=FALSE}
organism_raw <- myload_raw(organism_name)
```

```{r , echo=TRUE, eval=FALSE}
organism_raw |> slice_head(n = 3) |> as_tibble() |> glimpse()
```

## Step 6: Count Occurrences by State

```{r "number of occ in state", echo=TRUE, eval=FALSE}
organism_raw |> count(sci_name, obs_state)
```

## Step 7: Clean Taxonomic Ambiguity

```{r echo=TRUE, fig.width=6, fig.height=4, eval=FALSE}
radius_threshold <- 250000

unknowns <- organism_raw |> filter(sci_name == "Arachnocampa")
references <- organism_raw |> filter(sci_name == "Arachnocampa flava")

assign_if_close_to_flava <- function(lat, lon, flava_data, radius) {
  dists <- distHaversine(
    matrix(c(lon, lat), ncol = 2),
    matrix(c(flava_data$obs_lon, flava_data$obs_lat), ncol = 2)
  )
  if (any(dists <= radius)) "Arachnocampa flava" else "Arachnocampa"
}

unknowns_updated <- unknowns |> 
  rowwise() |> 
  mutate(sci_name = assign_if_close_to_flava(obs_lat, obs_lon, references, radius_threshold)) |> 
  ungroup()

organism <- organism_raw |> 
  filter(sci_name != "Arachnocampa") |> 
  bind_rows(unknowns_updated) |> 
  filter(!is.na(obs_lat), !is.na(obs_lon), !is.na(month),
         sci_name %in% c("Arachnocampa flava", "Arachnocampa tasmaniensis"))

mysave(organism_name, organism)
```

## Weather Station attach

```{r echo=TRUE, fig.width=6, fig.height=4, eval=FALSE}
attach_nearest_station(organism_name)
load("../data/glowworms.rda")
glimpse(head(glowworms, 2))
```
