---
title: "weather_data_reproduce"
format: html
---

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
source(here::here("data-raw/data_reproduce_utiles.R"))
```

## Step 1: Identify Top 3 Weather Stations per Organism

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
organism_names <- c("orchids", "gouldian_finch", "manta_rays", "glowworms")
load("../data/weather_stations.rda")
all_top_stations <- list()

for (org in organism_names) {
  load(paste0("../data/", org, ".rda"))
  org_data <- get(org)

  top_stations <- org_data %>%
    count(ws_id, sort = TRUE) %>%
    slice_max(n, n = 3) %>%
    left_join(weather_stations, by = c("ws_id")) %>%
    mutate(organism = org) %>%
    select(organism, ws_id)

  all_top_stations[[org]] <- top_stations
}

top_stations <- bind_rows(all_top_stations)
save(top_stations, file = "../data/top_stations.rda")
```

## Optional: Download Weather Data from GSOD

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
# Download and prepare data (optional)
weather_data_list <- lapply(top_stations$ws_id, function(stn) {
  get_GSOD(years = 2014:2024, station = stn)
})

top_weather <- bind_rows(weather_data_list) |> 
  transmute(
    ws_id = STNID,
    stn_lat = LATITUDE,
    stn_lon = LONGITUDE,
    date = YEARMODA,
    year = YEAR,
    month = MONTH,
    day = DAY,
    weekday = lubridate::wday(date, label = TRUE, abbr = FALSE,
                               week_start = getOption("lubridate.week.start", 1)),
    dayofyear = YDAY,
    temp = TEMP,
    min = MIN,
    max = MAX,
    dewp = DEWP,
    rh = RH,
    prcp = PRCP,
    rainy = ifelse(PRCP > 5, 1, 0),
    wind_speed = WDSP,
    max_speed = MXSPD
  )

# Save to data folder
assign("weather_data", top_weather)
save(weather_data, file = "../data/weather_data.rda")
```

## Conclusion

You now have a decade of daily weather observations for the most relevant weather stations associated with each organism in the ecotourism package. This can be used to study climate impacts, seasonal patterns, or ecological correlations with biodiversity data.
