---
title: "Manta Rays"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    toc-location: left
    fig-cap-location: top
    code-fold: false
    code-tools: true
    theme: flatly
    page-layout: full
editor: visual
---

---
vignette: >
  %\VignetteIndexEntry{Manta Rays}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Ensure the temporary library from R CMD check is visible (esp. on Windows)
libdir <- Sys.getenv("R_LIBS")
if (nzchar(libdir)) {
  parts <- strsplit(libdir, .Platform$path.sep, fixed = TRUE)[[1]]
  .libPaths(unique(c(parts, .libPaths())))
}

# now load your package
suppressPackageStartupMessages(library(ecotourism))
```

![Manta Rays at Dharavandhoo Thila (Black Pearl).\
Image credit: Shiyam ElkCloner – Own work. Licence: [CC BY-SA 3.0](https://creativecommons.org/licenses/by-sa/3.0/).\
File: Dharavandhoo Thila - Manata Black Pearl.JPG. Created: 27 May 2012. Uploaded: 27 May 2012.](image/manta_rays.jpg){fig-align="center" width="300"}

## Introduction

This vignette walks you through the process of downloading, cleaning, and analyzing occurrence data for **Manta Rays** in Australia. The data is sourced from the [Atlas of Living Australia (ALA)](https://en.wikipedia.org/wiki/Manta_ray), enriched with weather metadata, and visualized using custom functions.

------------------------------------------------------------------------

## Load Required Packages and Utilities

All helper functions are stored in `vignette-utils.R`:

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
#source(here::here("inst/vignette-helpers/vignette-utils.R"))
utils_path <- system.file("vignette-helpers", "vignette-utils.R", package = "ecotourism")
if (file.exists(utils_path)) {
  source(utils_path, encoding = "UTF-8")
}
```

```{r}
organism_name = "manta_rays"
organism_title = "Manta Rays"
```

------------------------------------------------------------------------

## Data Analysis

### Preview Dataset

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=TRUE}
data("manta_rays")
manta_rays |>  glimpse()
```

### Occurrence by State

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=TRUE}
manta_rays |> count(sci_name, obs_state)
```

------------------------------------------------------------------------

## Visualization

### Spatial Distribution Map

```{r echo=TRUE, eval=TRUE, fig.width=6, fig.height=4}
manta_rays |> plot_map1(legend = FALSE)
```

### Weekly Trends

```{r echo=TRUE, eval=TRUE, fig.width=6, fig.height=4}
manta_rays |> plot_bar_week(legend = FALSE)
```

### Monthly Trends

```{r echo=TRUE, eval=TRUE, fig.width=6, fig.height=4}
manta_rays |> plot_bar_month(legend = FALSE)
```

### Yearly Trends

```{r echo=TRUE, eval=TRUE, fig.width=6, fig.height=4}
manta_rays |> plot_bar_year(legend = FALSE)
```

------------------------------------------------------------------------

## Pattern Analysis

```{r echo=TRUE, eval=TRUE, fig.width=6, fig.height=4}
manta_rays |> plot_monthly_occurrences_by_state()
manta_rays |> plot_weekly_occurrences_by_state()
manta_rays |> plot_seasonal_occurrence_polar()
manta_rays |> plot_monthly_occurrence_map()
manta_rays |> plot_monthly_occurrence_map_by_state()
manta_rays |> plot_monthly_distribution_map()
```

------------------------------------------------------------------------

Work with Tourism dataset:

```{r, warning=FALSE, message=FALSE}
# make full tourism dataset
data("tourism_region")
data("tourism_quarterly")

tourism <- tourism_quarterly |>
  dplyr::left_join(tourism_region, by = "region_id") |> 
  dplyr::left_join(ecotourism::weather_stations, by = "ws_id")
```

```{r}
# load top station for organism
data("top_stations")

top_stations_manta_rays <- top_stations |> filter(organism == "manta_rays")
```

No we can find related tourism spots

```{r}
tourism |> filter(ws_id %in% top_stations_manta_rays$ws_id)
```

If you want to match quarter data use this code:

```{r}
# 1. Aggregate manta observations to quarter
manta_quarter <- manta_rays %>%
  mutate(quarter = floor_date(date, "quarter")) %>%
  count(quarter, name = "manta_sightings")

# 2. Make sure tourism$quarter is also floor_date
tourism_clean <- tourism %>%
  mutate(quarter = floor_date(quarter, "quarter"))

# 3. Left join: keep all manta quarters, match tourism by quarter
manta_tourism <- manta_quarter %>%
  left_join(tourism_clean, by = "quarter")
```
