---
title: "Tourism data"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    toc-location: left
    fig-cap-location: top
    code-fold: false
    code-tools: true
    theme: flatly
    page-layout: full
editor: visual
---

---
vignette: >
  %\VignetteIndexEntry{tourism data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

::: callout-note
### What this vignette shows

A quick tour of the **ecotourism** package's tourism datasets, how to join regional metadata and quarterly counts, and (optionally) how to enrich rows with the nearest weather station information.
:::

## Setup

```{r}
#| eval: false
#| message: true
#| warning: false
#| include: true
# Core tidy tools
library(dplyr)
library(ggplot2)

# Spatial & basemap (suggests)
# If these aren't installed, install with: pak::pak(c("sf","rnaturalearth","rnaturalearthdata"))
library(sf)
library(rnaturalearth)

# Package data
library(ecotourism)
```


## Data overview

List included objects and short descriptions:

```{r}
#| eval: false
#| message: true
#| warning: false
#| include: true
load("../data/tourism_quarterly.rda")
load("../data/tourism_region.rda")
```

Quick peeks at the two tourism datasets used here:

```{r}
#| eval: false
#| message: true
#| warning: false
#| include: true
dplyr::glimpse(tourism_region)
dplyr::glimpse(tourism_quarterly)
```

## Join regional metadata to quarterly counts

We keep only rows with a valid region name after the join.

```{r}
#| eval: false
#| message: true
#| warning: false
#| include: true
tourism <- tourism_quarterly |>
  dplyr::left_join(tourism_region, by = "region_id") |>
  dplyr::filter(!is.na(region))

dplyr::glimpse(tourism)
```

::: callout-tip
If you only need a subset of columns, select them first to keep memory small, for example: `select(region_id, quarter, visitors, lon, lat)`.
:::

## Optional: add nearest weather station metadata

If your quarterly table contains a `ws_id`, you can join station attributes:

```{r}
#| eval: false
#| message: false
#| warning: false
#| include: true
  tourism <- tourism |>
    dplyr::left_join(ecotourism::weather_stations, by = "ws_id")
```


## A quick map

We’ll plot tourism site coordinates and, when available, their weather stations.

```{r}
#| eval: false
#| message: true
#| warning: false
#| include: true
# Keep rows with geographic coordinates
spots <- tourism |> dplyr::filter(!is.na(lon), !is.na(lat))

# Convert to sf (WGS84)
spots_sf <- sf::st_as_sf(spots, coords = c("lon","lat"), crs = 4326, remove = FALSE)

# Australia basemap
aus <- rnaturalearth::ne_states(country = "australia", returnclass = "sf")

# Plot
ggplot() +
  geom_sf(data = aus, fill = "grey95", color = "grey60", linewidth = 0.3) +
  geom_sf(data = spots_sf, color = "#2C7FB8", alpha = 0.8, size = 1.6) +
  coord_sf(xlim = c(110, 155), ylim = c(-45, -10), expand = FALSE) +
  labs(title = "Tourism Locations in Australia", x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major = element_line(color = "grey90"))
```

If station coordinates `stn_lon` and `stn_lat` are present, overlay them too:

```{r}
#| eval: false
#| message: true
#| warning: false
#| include: true
if (all(c("stn_lon","stn_lat") %in% names(tourism))) {
  ggplot() +
    geom_sf(data = aus, fill = "grey95", color = "grey60", linewidth = 0.3) +
    geom_point(data = tourism, aes(x = lon, y = lat),
               alpha = 0.8, size = 2) +
    geom_point(data = tourism, aes(x = stn_lon, y = stn_lat),
               shape = 17, size = 1.5) +
    coord_sf(xlim = c(110, 155), ylim = c(-45, -10), expand = FALSE) +
    labs(title = "Tourism Locations and Weather Stations",
         x = NULL, y = NULL) +
    theme_minimal(base_size = 12) +
    theme(panel.grid.major = element_line(color = "grey90"))
}
```

## Simple time-series slice

Plot quarterly visitors for one example region (pick the first with data).

```{r}
#| eval: false
#| message: true
#| warning: false
#| include: true
if (all(c("region","quarter","trips") %in% names(tourism))) {
  example_region <- tourism |>
    dplyr::filter(!is.na(trips)) |>
    dplyr::count(region, wt = trips, name = "total", sort = TRUE) |>
    dplyr::slice(1) |>
    dplyr::pull(region)

  tourism |>
    dplyr::filter(region == example_region) |>
    ggplot(aes(x = quarter, y = trips)) +
    geom_line() +
    labs(title = paste0("Quarterly trips – ", example_region),
         x = NULL, y = "trips") +
    theme_minimal()
}
```

