---
title: "Tourism data"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    toc-location: left
    code-fold: false
    code-tools: true
    theme: flatly
    page-layout: full
editor: visual
vignette: >
  %\VignetteIndexEntry{tourism data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
libdir <- Sys.getenv("R_LIBS")
if (nzchar(libdir)) {
  parts <- strsplit(libdir, .Platform$path.sep, fixed = TRUE)[[1]]
  .libPaths(unique(c(parts, .libPaths())))
}

suppressPackageStartupMessages(library(ecotourism))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(sf))
```

The `ecotourism` package contains complementary tourism datasets at quarterly tourism-region and monthly SA2 resolutions. This vignette introduces the monthly stopover reason and activity datasets, their code-name lookup tables, and the geographic metadata used to map them.

## Data overview

`tourism_reason` contains monthly domestic overnight-trip estimates classified by the reason for the stopover.

```{r, message=FALSE}
data("tourism_reason")
glimpse(tourism_reason)
```

`tourism_activity` contains the corresponding estimates classified by activity.

```{r, message=FALSE}
data("tourism_activity")
glimpse(tourism_activity)
```

Both datasets cover 2014--2024. Trip estimates are expressed in thousands, and `n_sample` gives the survey sample supporting each estimate. Rows with zero `trips` or zero `n_sample` have been removed.

## Add descriptive category names

The compact integer category fields can be decoded with `tourism_reason_name` and `tourism_activity_name`.

```{r, message=FALSE}
data("tourism_reason_name")
data("tourism_activity_name")

tourism_reason_name
tourism_activity_name
```

Join the lookup table appropriate to the detailed dataset:

```{r, message=FALSE}
reason_named <- tourism_reason |>
  left_join(tourism_reason_name, by = "stop_reason")

activity_named <- tourism_activity |>
  left_join(tourism_activity_name, by = "stop_activity")

reason_named |>
  count(year, name, wt = trips, name = "trips_thousands") |>
  arrange(year, desc(trips_thousands)) |>
  head()
```

## Add tourism-region metadata

The detailed datasets share `sa2_code_tra` with `tourism_region`. The packaged region table stores this identifier as character, so harmonise the type before joining.

```{r, message=FALSE}
data("tourism_region")

region_lookup <- tourism_region |>
  mutate(sa2_code_tra = as.integer(sa2_code_tra)) |>
  distinct(sa2_code_tra, state, region, sa2, lon, lat, ws_id)

reason_with_region <- reason_named |>
  left_join(region_lookup, by = "sa2_code_tra")

reason_with_region |>
  select(year, month, state, region, sa2, name, trips, n_sample) |>
  head()
```

The same `region_lookup` can be joined to `tourism_activity`. The `ws_id` column can subsequently be joined to `weather_stations` when weather context is needed.

## SA2 polygons and representative points

`oz_sa2` provides simplified 2021 SA2 polygons for every TRA code represented by the detailed tourism data. Its `lat` and `lon` columns provide one representative interior point per SA2 when point geometry is more convenient.

```{r, message=FALSE}
data("oz_sa2")
glimpse(oz_sa2)
```

Join an SA2 summary directly to the polygon table for mapping:

```{r, message=FALSE}
reason_by_sa2 <- tourism_reason |>
  summarise(
    trips_thousands = sum(trips),
    .by = sa2_code_tra
  )

reason_map <- oz_sa2 |>
  left_join(reason_by_sa2, by = "sa2_code_tra")

reason_map |>
  select(sa2_code_tra, lat, lon, trips_thousands) |>
  head()
```

## Recover omitted zero rows

Use `tidyr::complete()` to restore the full monthly panels; combinations absent from the packaged data are assigned zero `trips` and `n_sample`.

```{r, message=FALSE}
all_sa2 <- union(tourism_reason$sa2_code_tra, tourism_activity$sa2_code_tra)
months <- factor(month.name, levels = month.name, ordered = TRUE)

tourism_reason_full <- tourism_reason |>
  tidyr::complete(year = 2014:2024, month = months, sa2_code_tra = all_sa2,
                  stop_reason = tourism_reason_name$stop_reason,
                  fill = list(trips = 0, n_sample = 0))

tourism_activity_full <- tourism_activity |>
  tidyr::complete(year = 2014:2024, month = months, sa2_code_tra = all_sa2,
                  stop_activity = tourism_activity_name$stop_activity,
                  fill = list(trips = 0, n_sample = 0))
```