---
title: "Tourism data"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    toc-location: left
    fig-cap-location: top
    code-fold: false
    code-tools: true
    theme: flatly
    page-layout: full
editor: visual
---

---
vignette: >
  %\VignetteIndexEntry{tourism data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Ensure the temporary library from R CMD check is visible (esp. on Windows)
libdir <- Sys.getenv("R_LIBS")
if (nzchar(libdir)) {
  parts <- strsplit(libdir, .Platform$path.sep, fixed = TRUE)[[1]]
  .libPaths(unique(c(parts, .libPaths())))
}

# now load your package
suppressPackageStartupMessages(library(ecotourism))
```

::: callout-note
### What this vignette shows

A quick tour of the **ecotourism** package's tourism datasets, how to join regional metadata and quarterly counts, and (optionally) how to enrich rows with the nearest weather station information.
:::

## Setup

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
#source(here::here("inst/vignette-helpers/vignette-utils.R"))
utils_path <- system.file("vignette-helpers", "vignette-utils.R", package = "ecotourism")
if (file.exists(utils_path)) {
  source(utils_path, encoding = "UTF-8")
}
```

## Data overview

List included objects and short descriptions:

```{r}
data("tourism_region")
data("tourism_quarterly")
```

Quick peeks at the two tourism datasets used here:

```{r}
#| message: true
#| warning: false
#| include: true
dplyr::glimpse(tourism_region)
dplyr::glimpse(tourism_quarterly)
```

## Join regional metadata to quarterly counts

We keep only rows with a valid region name after the join.

```{r}
#| message: true
#| warning: false
#| include: true
tourism <- tourism_quarterly |>
  dplyr::left_join(tourism_region, by = "region_id") |>
  dplyr::filter(!is.na(region))

dplyr::glimpse(tourism)
```

::: callout-tip
If you only need a subset of columns, select them first to keep memory small, for example: `select(region_id, quarter, visitors, lon, lat)`.
:::

## Optional: add nearest weather station metadata

If your quarterly table contains a `ws_id`, you can join station attributes:

```{r}
#| message: false
#| warning: false
#| include: true
  tourism <- tourism |>
    dplyr::left_join(ecotourism::weather_stations, by = "ws_id")
```

## A quick map

We’ll plot tourism site coordinates and, when available, their weather stations.

```{r}
#| message: true
#| warning: false
#| include: true
# Keep rows with geographic coordinates
spots <- tourism |> dplyr::filter(!is.na(lon), !is.na(lat))

aus <- oz_lga

# Plot
ggplot() +
  geom_sf(data = aus, fill = "grey95", color = "grey60", linewidth = 0.3) +
  geom_point(data = spots, aes(x = lon, y = lat),
               alpha = 0.8, size = 2) +
  coord_sf(xlim = c(110, 155), ylim = c(-45, -10), expand = FALSE) +
  labs(title = "Tourism Locations in Australia", x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major = element_line(color = "grey90"))
```

If station coordinates `stn_lon` and `stn_lat` are present, overlay them too:

```{r}
#| message: true
#| warning: false
#| include: true
if (all(c("stn_lon","stn_lat") %in% names(tourism))) {
  ggplot() +
    geom_sf(data = aus, fill = "grey95", color = "grey60", linewidth = 0.3) +
    geom_point(data = tourism, aes(x = lon, y = lat),
               alpha = 0.5, size = 2, color = "blue") +
    geom_point(data = tourism, aes(x = stn_lon, y = stn_lat),
               shape = 17, size = 1.5, alpha = 0.5, color = "red") +
    coord_sf(xlim = c(110, 155), ylim = c(-45, -10), expand = FALSE) +
    labs(title = "Tourism Locations and Weather Stations",
         x = NULL, y = NULL) +
    theme_minimal(base_size = 12) +
    theme(panel.grid.major = element_line(color = "grey90"))
}
```

## Simple time-series slice

Plot quarterly visitors for one example region (pick the first with data).

```{r}
#| message: true
#| warning: false
#| include: true
if (all(c("region","quarter","trips") %in% names(tourism))) {
  example_region <- tourism |>
    dplyr::filter(!is.na(trips)) |>
    dplyr::count(region, wt = trips, name = "total", sort = TRUE) |>
    dplyr::slice(1) |>
    dplyr::pull(region)

  tourism |>
    dplyr::filter(region == example_region) |>
    ggplot(aes(x = quarter, y = trips)) +
    geom_line() +
    labs(title = paste0("Quarterly trips – ", example_region),
         x = NULL, y = "trips") +
    theme_minimal()
}
```
