---
title: "Weather Data"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    toc-location: left
    fig-cap-location: top
    code-fold: false
    code-tools: true
    theme: flatly
    page-layout: full
editor: visual
---

---
vignette: >
  %\VignetteIndexEntry{Weather Data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Ensure the temporary library from R CMD check is visible (esp. on Windows)
libdir <- Sys.getenv("R_LIBS")
if (nzchar(libdir)) {
  parts <- strsplit(libdir, .Platform$path.sep, fixed = TRUE)[[1]]
  .libPaths(unique(c(parts, .libPaths())))
}

# now load your package
suppressPackageStartupMessages(library(ecotourism))
```


## Introduction

This vignette explains how to retrieve and organize **weather data** for ecotourism analysis in Australia.\
We identify the top 3 weather stations associated with each focal organism and optionally download 10 years of daily weather data (2014–2024) for each.

------------------------------------------------------------------------

## Load Utilities

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
source("vignette-utils.R")
```

------------------------------------------------------------------------

## Step 1: Identify Top 3 Weather Stations per Organism

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
organism_names <- c("orchids", "gouldian_finch", "manta_rays", "glowworms")
load("../data/weather_stations.rda")
all_top_stations <- list()

for (org in organism_names) {
  load(paste0("../data/", org, ".rda"))
  org_data <- get(org)

  top_stations <- org_data %>%
    count(ws_id, sort = TRUE) %>%
    slice_max(n, n = 3) %>%
    left_join(weather_stations, by = c("ws_id")) %>%
    mutate(organism = org) %>%
    select(organism, ws_id)

  all_top_stations[[org]] <- top_stations
}

top_stations <- bind_rows(all_top_stations)
save(top_stations, file = "../data/top_stations.rda")
```

------------------------------------------------------------------------

## Top 3 Stations for Each Organism

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
load("../data/weather_stations.rda")
load("../data/top_stations.rda")
df <- left_join(top_stations, weather_stations, by = "ws_id")
df
```

------------------------------------------------------------------------

## Map of Top Weather Stations

```{r echo=TRUE, eval=FALSE, fig.width=6, fig.height=4}
aus_map <- ne_states(country = "Australia", returnclass = "sf")
ggplot() +
  geom_sf(data = aus_map, fill = "gray95", color = "gray50") +
  geom_point(data = df, aes(x = stn_lon, y = stn_lat, color = organism), shape = 2, alpha = 0.8, size = 3) +
  coord_sf(xlim = c(110, 155), ylim = c(-45, -10)) +
  theme_minimal()
```

------------------------------------------------------------------------

## Optional: Download Weather Data from GSOD

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
# Download and prepare data (optional)
weather_data_list <- lapply(top_stations$ws_id, function(stn) {
  get_GSOD(years = 2014:2024, station = stn)
})

top_weather <- bind_rows(weather_data_list) |> 
  transmute(
    ws_id = STNID,
    stn_lat = LATITUDE,
    stn_lon = LONGITUDE,
    date = YEARMODA,
    year = YEAR,
    month = MONTH,
    day = DAY,
    weekday = lubridate::wday(date, label = TRUE, abbr = FALSE,
                               week_start = getOption("lubridate.week.start", 1)),
    dayofyear = YDAY,
    temp = TEMP,
    min = MIN,
    max = MAX,
    dewp = DEWP,
    rh = RH,
    prcp = PRCP,
    rainy = ifelse(PRCP > 5, 1, 0),
    wind_speed = WDSP,
    max_speed = MXSPD
  )

# Save to data folder
assign("weather_data", top_weather)
save(weather_data, file = "../data/weather_data.rda")
```

------------------------------------------------------------------------

## Preview of Final Weather Dataset

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
load("../data/weather_data.rda")
weather_data |> glimpse()
```

------------------------------------------------------------------------

## Conclusion

You now have a decade of daily weather observations for the most relevant weather stations associated with each organism in the ecotourism package. This can be used to study climate impacts, seasonal patterns, or ecological correlations with biodiversity data.

## Final Output

This is the glimps of your final data :

```{r, echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
load("../data/weather_data.rda")
load("../data/top_stations.rda")
top_stations
weather_data |> glimpse()
```

```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ecotourism)
data("top_stations")
data("weather_data")
top_stations
weather_data |> glimpse()
```
