---
title: "weather_data_2024"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{weather_data_2024}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#### Download Weather Data for Top 3 Stations

```{r echo=FALSE, message=FALSE, warning=FALSE}
# load required packages
library(GSODR)
library(dplyr)
library(ggplot2)
library(tidyverse)
```
##### Top 3 Station for **Orchids**
```{r echo=FALSE}
# find top 3 station 

load("../data/orchids_2024_ws.rda")
load("../data/weather_stations.rda")

orchids_station_counts <- orchids_2024_ws %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) |> 
  distinct()

head(orchids_station_counts, n = 3)
```

```{r echo=FALSE, eval=FALSE}
orchids_top_stations <- orchids_station_counts |> slice_max(n, n=3)

orchids_weather_data_list <- lapply(orchids_top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine if needed
orchids_top_weather <- dplyr::bind_rows(orchids_weather_data_list)

usethis::use_data(orchids_top_weather)
```

##### Top 3 Station for **Gouldian finch**
```{r echo=FALSE}
organism <- "gouldian_finch"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)

```

##### Top 3 Station for **Hooded parrot**
```{r echo=FALSE}
organism <- "hooded_parrot"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)

```

##### Top 3 Station for **Golden-shouldered parrot**
```{r echo=FALSE}
organism <- "golden_shouldered_parrot"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)

```

##### Top 3 Station for **Southern black-throated finch**
```{r echo=FALSE}
organism <- "southern_black_throated_finch"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)

```

##### Top 3 Station for **Humpback whale**
```{r echo=FALSE}
organism <- "whales"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)

```


##### Top 3 Station for **Whale Sharks**
```{r echo=FALSE}
organism <- "whale_sharks"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)

```

##### Top 3 Station for **Koala**
```{r echo=FALSE}
organism <- "koala"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)

```
```Warning :
This station, 947860-99999, only provides data for years 1996 to 2022.
Please send a request that falls within these years.
```


##### Top 3 Station for **Platypus**
```{r echo=FALSE}
organism <- "platypus"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)
```

##### Top 3 Station for **Wombat**
```{r echo=FALSE}
organism <- "wombat"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)
```


##### Top 3 Station for **Glowworms**
```{r echo=FALSE}
organism <- "glowworms"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)
```
```
Warning :
This station, 949630-99999, only provides data for years 1968 to 2020.
Please send a request that falls within these years.
```

##### Top 3 Station for **Penguins**
```{r echo=FALSE}
organism <- "penguins"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)
```

##### Top 3 Station for **Giant cuttlefish**
```{r echo=FALSE}
organism <- "giant_cuttlefish"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)
```

```
Warning :
This station, 957790-99999, only provides data for years 2003 to 2023.
Please send a request that falls within these years.
```
##### Top 3 Station for **Manta Rays**
```{r echo=FALSE}
organism <- "manta_rays"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)
```

##### Top 3 Station for **Salt Water Crocodiles**
```{r echo=FALSE}
organism <- "saltwater_crocodiles"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)
```

##### Top 3 Station for **Thorny Devil**
```{r echo=FALSE}
organism <- "thorny_devil"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)
```

##### Top 3 Station for **Sea Turtles**
```{r echo=FALSE}
organism <- "sea_turtles"
# Load required data
load(paste0("../data/", organism, "_2024_ws.rda"))
load("../data/weather_stations.rda")

# Identify top 3 stations
station_counts <- get(paste0(organism, "_2024_ws")) %>%
  count(weather_station_id, sort = TRUE) %>%
  left_join(weather_stations, by = c("weather_station_id" = "stnid")) %>%
  distinct()

head(station_counts, n = 3)

```

```{r echo=FALSE, eval=FALSE}
# Get weather data for top 3 stations (optional)
top_stations <- station_counts |> slice_max(n, n = 3)

weather_data_list <- lapply(top_stations$weather_station_id, function(stn) {
  get_GSOD(years = 2024, station = stn)
})

# Combine and save
top_weather <- dplyr::bind_rows(weather_data_list)

# Dynamically assign it to the global environment under a specific name
obj_name <- paste0(organism, "_top_weather")
file_path <- file.path("../data", paste0(obj_name, ".rda"))
assign(obj_name, top_weather)

save(list = obj_name, file = file_path)
```
