---
title: "Australia Weather Station"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    toc-location: left
    fig-cap-location: top
    code-fold: false
    code-tools: true
    theme: flatly
    page-layout: full
editor: visual
---

---
vignette: >
  %\VignetteIndexEntry{Australia Weather Station}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Ensure the temporary library from R CMD check is visible (esp. on Windows)
libdir <- Sys.getenv("R_LIBS")
if (nzchar(libdir)) {
  parts <- strsplit(libdir, .Platform$path.sep, fixed = TRUE)[[1]]
  .libPaths(unique(c(parts, .libPaths())))
}

# now load your package
suppressPackageStartupMessages(library(ecotourism))
```

## Introduction

This vignette demonstrates how to download, process, and visualize the locations of weather stations across **Australia** using the `GSODR` package. These stations are key to understanding long-term climate patterns in ecotourism studies.

------------------------------------------------------------------------

## Load Utilities

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
source("vignette-utils.R")
```

------------------------------------------------------------------------

## Step 1: Download and Clean Weather Station Data

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
# Download and filter weather stations for Australia
stations_all <- get_inventory()

weather_stations <- stations_all |>
  filter(CTRY == "AS") |>
  select(STNID, NAME, LAT, LON) |> 
  rename(stn_lat = LAT, stn_lon = LON, ws_id = STNID, stname = NAME) |> 
  distinct(stn_lat, stn_lon, .keep_all = TRUE) |> 
  filter(stn_lat >= -44, stn_lat <= -8, stn_lon >= 110, stn_lon <= 160) |> 
  reverse_geocode(lat = stn_lat, long = stn_lon, method = "osm")
```

------------------------------------------------------------------------

## Step 2: Extract State Information from Addresses

```{r eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
weather_stations <- weather_stations %>%
  rowwise() %>%
  mutate(parsed = list(parse_address(address))) %>%
  unnest(parsed) %>%
  ungroup()
```

------------------------------------------------------------------------

## Step 3: Save Data for Later Use

```{r eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
usethis::use_data(weather_stations, overwrite = TRUE)
```

------------------------------------------------------------------------

## Weather Station Distribution by State

```{r eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
load("../data/weather_stations.rda")
weather_stations |> count(stn_state, sort = TRUE) |> kable()
```

------------------------------------------------------------------------

## Preview of Weather Station Data

```{r eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
load("../data/weather_stations.rda")
weather_stations |> glimpse()
```

```{r eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ecotourism)
data("weather_stations")
weather_stations |> glimpse()
```

------------------------------------------------------------------------

## Visualization: Map of Weather Stations

```{r eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
aus_map <- ne_states(country = "Australia", returnclass = "sf")

ggplot() +
  geom_sf(data = aus_map, fill = "gray95", color = "gray50") +
  geom_point(data = weather_stations, 
             aes(x = stn_lon, y = stn_lat, colour = stn_state),
             alpha = 0.6, size = 2) +
  coord_sf(xlim = c(115, 155), ylim = c(-45, -10)) +
  labs(
    title = "Weather Stations in Australia",
    x = "Longitude",
    y = "Latitude",
    colour = "State"
  ) +
  theme_minimal()
```

------------------------------------------------------------------------

## Conclusion

You now have a cleaned, geocoded, and mapped dataset of weather stations across Australia. This can be used to join with ecological data to analyze regional weather trends in biodiversity and tourism patterns.
