---
title: "Australia Weather Station"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    toc-location: left
    fig-cap-location: top
    code-fold: false
    code-tools: true
    theme: flatly
    page-layout: full
editor: visual
---

---
vignette: >
  %\VignetteIndexEntry{Australia Weather Station}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Ensure the temporary library from R CMD check is visible (esp. on Windows)
libdir <- Sys.getenv("R_LIBS")
if (nzchar(libdir)) {
  parts <- strsplit(libdir, .Platform$path.sep, fixed = TRUE)[[1]]
  .libPaths(unique(c(parts, .libPaths())))
}

# now load your package
suppressPackageStartupMessages(library(ecotourism))
```

## Introduction

This vignette demonstrates how to download, process, and visualize the locations of weather stations across **Australia** using the `GSODR` package. These stations are key to understanding long-term climate patterns in ecotourism studies.

------------------------------------------------------------------------

## Load Utilities

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
#source(here::here("inst/vignette-helpers/vignette-utils.R"))
utils_path <- system.file("vignette-helpers", "vignette-utils.R", package = "ecotourism")
if (file.exists(utils_path)) {
  source(utils_path, encoding = "UTF-8")
}
```

------------------------------------------------------------------------

## Weather Station Distribution by State

```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(ecotourism)
data("weather_stations")
weather_stations |> count(stn_state, sort = TRUE) 
```

------------------------------------------------------------------------

## Preview of Weather Station Data

```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ecotourism)
data("weather_stations")
weather_stations |> glimpse()
```

------------------------------------------------------------------------

## Visualization: Map of Weather Stations

```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
data(oz_lga)
aus_map <- oz_lga

ggplot() +
  geom_sf(data = aus_map, fill = "gray95", color = "gray50") +
  geom_point(data = weather_stations, 
             aes(x = stn_lon, y = stn_lat, colour = stn_state),
             alpha = 0.6, size = 2) +
  coord_sf(xlim = c(115, 155), ylim = c(-45, -10)) +
  labs(
    title = "Weather Stations in Australia",
    x = "Longitude",
    y = "Latitude",
    colour = "State"
  ) +
  theme_minimal()
```

------------------------------------------------------------------------

## Conclusion

You now have a cleaned, geocoded, and mapped dataset of weather stations across Australia. This can be used to join with ecological data to analyze regional weather trends in biodiversity and tourism patterns.
