---
title: "Weather Data"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    toc-location: left
    fig-cap-location: top
    code-fold: false
    code-tools: true
    theme: flatly
    page-layout: full
editor: visual
---

---
vignette: >
  %\VignetteIndexEntry{Weather Data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Ensure the temporary library from R CMD check is visible (esp. on Windows)
libdir <- Sys.getenv("R_LIBS")
if (nzchar(libdir)) {
  parts <- strsplit(libdir, .Platform$path.sep, fixed = TRUE)[[1]]
  .libPaths(unique(c(parts, .libPaths())))
}

# now load your package
suppressPackageStartupMessages(library(ecotourism))
```

## Introduction

This vignette explains how to retrieve and organize **weather data** for ecotourism analysis in Australia.\
We identify the top 3 weather stations associated with each focal organism and optionally download 10 years of daily weather data (2014–2024) for each.

------------------------------------------------------------------------

## Load Utilities

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
#source(here::here("inst/vignette-helpers/vignette-utils.R"))
utils_path <- system.file("vignette-helpers", "vignette-utils.R", package = "ecotourism")
if (file.exists(utils_path)) {
  source(utils_path, encoding = "UTF-8")
}
```

------------------------------------------------------------------------

## Top 3 Stations for Each Organism

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=TRUE}
data("top_stations")
data("weather_stations")
df <- left_join(top_stations, weather_stations, by = "ws_id")
df |> glimpse()
```

------------------------------------------------------------------------

## Map of Top Weather Stations

```{r echo=TRUE, eval=FALSE, fig.width=6, fig.height=4}
aus_map <- ne_states(country = "Australia", returnclass = "sf")
ggplot() +
  geom_sf(data = aus_map, fill = "gray95", color = "gray50") +
  geom_point(data = df, aes(x = stn_lon, y = stn_lat, color = organism), shape = 2, alpha = 0.8, size = 3) +
  coord_sf(xlim = c(110, 155), ylim = c(-45, -10)) +
  theme_minimal()
```

------------------------------------------------------------------------

## Final Weather Data-set

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=TRUE}
data("weather_data")
weather_data |> glimpse()
```
