---
title: "Weather Data"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    toc-location: left
    fig-cap-location: top
    code-fold: false
    code-tools: true
    theme: flatly
    page-layout: full
editor: visual
---

---
vignette: >
  %\VignetteIndexEntry{Weather Data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Ensure the temporary library from R CMD check is visible (esp. on Windows)
libdir <- Sys.getenv("R_LIBS")
if (nzchar(libdir)) {
  parts <- strsplit(libdir, .Platform$path.sep, fixed = TRUE)[[1]]
  .libPaths(unique(c(parts, .libPaths())))
}

# now load your package
suppressPackageStartupMessages(library(ecotourism))
```

## Introduction

This vignette explains how to retrieve and organize **weather data** for ecotourism analysis in Australia.\
We identify the top 3 weather stations associated with each focal organism and optionally download 10 years of daily weather data (2014–2024) for each.

------------------------------------------------------------------------

This is the glimpse of your data :

```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ecotourism)

data("top_stations")
top_stations |> glimpse()
```

```{r, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
data("weather_stations")
weather_stations |> glimpse()
```

------------------------------------------------------------------------

## Map of Top Weather Stations

```{r echo=TRUE, eval=FALSE, fig.width=6, fig.height=4}

top_stations |> left_join(weather_stations) |> 
ggplot() +
  geom_sf(data = oz_lga) +
  geom_point(aes(x = stn_lon, y = stn_lat, color = organism), shape = 17, size = 3)
  theme_minimal()
```

------------------------------------------------------------------------

## Final Weather Data-set

```{r, echo=TRUE, message=FALSE, warning=FALSE, eval=TRUE}
data("weather_data")
weather_data |> glimpse()
```
